Exploring Alternative Routes Using Multipath TCP

### 2017-06-04 paper/presentation.tex
paper: git rid of out of date comments
presentation: updated related work from new changes

    \begin{itemize}
    \item Putting together, problem: unmodified applications can't take
      advantage of detour routing solutions to Internet routing inefficiencies.
    \item A solution could be a system which uses and aggregates alternative
      routes for MPTCP connections.
    \item Contribution 1 - a mechanism for adding new routes to MPTCP
      connections which are single-homed, with no modification to applications.
    \item Contribution 2 - experimental validation that the mechanism works,
      even at higher throughputs
    \end{itemize}

@@

\section{Related Work}
\SectionHead{}{
  \textbf{Transition:} Now in order to appreciate the place of this thesis,
  let's examine some related work. The topic includes detour routing, overlay
  routing, source routing, and multipath transport.
}

@@

  \begin{itemize}
  \item Rather than use only one detour, create an overlay network
  \item Overlay nodes use the Internet as their ``link layer''
    \note[item]{In Intro, discussed Detour study. RON is something of a
      spiritual successor. Rather than simple detours, use overlay networks.}
  \item Routing performed at each node using measured link characteristics
    \note[item]{Each node is an Internet host. Constantly monitoring link
      characteristics of network, updating routing tables with performance
      metrics.}
  \item Several studies based on RON:
    \note[item]{RON achieved good results, was testbed for additional studies.}
    \begin{itemize}
    \item Redundant multipath routing\mycite{andersen2003best}
      \note[item]{One study compared two routing strategies: adaptively choosing
      the best path, or redundantly using multiple paths. No aggregation.}
  \item ``Biologically inspired'' multipath routing
    \mycite{leibnitz2006biologically}
    \note[item]{Another - biologically inspired. Choose strongest path as
      primary, have several backups, involves differential equations and
      randomness.}
  \item mTCP \mycite{zhang2004transport}
    \note[item]{mTCP: most important. A modification of TCP before
      MPTCP existed. Built on RON + userspace TCP stack. Did not have MPTCP
      compatibility concerns - routed segments across different paths. Worked on
      single-homed devices due to RON}
    \end{itemize}
  \end{itemize}

@@

\begin{frame}
  \frametitle{Other Related Items}

  \note[item]{Outside of overlay networks, several related pieces of work, both
    in research and in the standards bodies.}

  \begin{itemize}
  \item Source routing (IPv4) and segment routing (IPv6)
    \note[item]{Notable - it is possible to specify intermediaries in both IPv4
      and IPv6. For security reasons these aren't respected. Standards exist
      mainly for use in internal networks.}
  \item Binder: aggregating bandwidth of several Internet gateways
    \mycite{binder}
    \note[item]{One use of MPTCP that is very similar was called Binder. A rural
      town was away from any ISP. Had several potential gateways far away, but
      low bandwidth and not good reliability. Established connections across
      each gateway, and created an OpenVPN connection to an Internet-side
      server, with subflow across each.}
  \item SOSR: Scalable One-hop Source Routing \mycite{gummadi2004improving}
    \note[item]{Work inspired by RON,
      but trying to simplify. This work simply tunneled connections across a
      single hop, using a NAT approach similar to ours. Found they got most of
      RON's benefit without the full overlay system.}
  \item IETF MPTCP proxying draft
    \note[item]{Similar because allows single-homed devices to use MPTCP.
      However this approach expects non MPTCP server, and is ``transparent''
      proxy.}
  \end{itemize}

>>>>

    \begin{itemize}
    \item Putting together, problem: unmodified applications can't take
      advantage of detour routing solutions to Internet routing inefficiencies.
    \item A solution could be a system which uses and aggregates alternative
      routes for MPTCP connections.
    \item Read contributions
    \end{itemize}

@@

\section{Related Work}
\SectionHead{}{

  \textbf{Transition:} Now in order to appreciate the place of this thesis,
  let's examine some related work. While there is a lot of related work, we will
  restrict this discussion to three major points: overlay networks (which
  operate at the network level), one very similar transport-layer approach, and
  some application level approaches.

@@

  \begin{itemize}
  \item Rather than use only one detour, create an overlay network
  \item Overlay nodes use the Internet as their ``link layer''
    \note[item]{In Intro, discussed Detour study. RON is something of a
      spiritual successor. Rather than simple detours, use overlay networks.}
  \item Routing performed at each node using measured link characteristics
    \note[item]{Each node is an Internet host. Constantly monitoring link
      characteristics of network, updating routing tables with performance
      metrics.}
  \item Several studies based on RON:
    \note[item]{RON achieved good results, was testbed for additional studies.}
    \begin{itemize}
    \item Redundant multipath routing\mycite{andersen2003best}
      \note[item]{One study compared two routing strategies: adaptively choosing
      the best path, or redundantly using multiple paths. No aggregation.}
  \item ``Biologically inspired'' multipath routing
    \mycite{leibnitz2006biologically}
    \note[item]{Another - biologically inspired. Choose strongest path as
      primary, have several backups, involves differential equations and
      randomness.}
  \item mTCP \mycite{zhang2004transport}
    \note[item]{mTCP: most important. A modification of TCP before
      MPTCP existed. Built on RON + userspace TCP stack. Did not have MPTCP
      compatibility concerns - routed segments across different paths. Worked on
      single-homed devices due to RON. Unmodified applications may not use.}
    \end{itemize}
  \end{itemize}

@@

\begin{frame}
  \frametitle{Application Layer}

  \note[item]{Outside of network-layer approaches, lots of app level approaches
    to overlays, detours, and bandwidth aggregation.}
  \note[item]{Since none of the net-layer approaches can be applied to
    unmodified apps, apps designed around overlays/detours are the only ones
    which can be used.}

  \begin{itemize}
    \item Gnutella\mycite{adar2000free}
      \note[item]{\textbf{Gnutella} was a file sharing network. It allowed
        queries and responses to be forwarded from peer to peer as an overlay.
        But actual file exchanges were single path.}
      \begin{itemize}
      \item Requests forwarded via overlay network
      \item Content exchanged via single path
      \end{itemize}
    \item BitTorrent\mycite{cohen2003incentives}
      \note[item]{\textbf{BitTorrent} is a protocol for file sharing which
        allows content to be exchanged along the overlay. Each peer shares with
        many other peers. Somewhat similar to detour routing - gets around
        routing inefficiencies by simply using more paths}
      \begin{itemize}
      \item Pieces of content exchanged between many pairs of peers
      \item Multiple paths simulate detour routing
      \end{itemize}
    \item HTTP Range Requests\mycite{kaspar2010enhancing}
      \note[item]{\textbf{HTTP Range Requests} allow an app to request byte
        ranges of a file. Kaspar et al demonstrate a mechanism for downloading
        or streaming a file using range requests on multiple network interfaces}
      \begin{itemize}
      \item Range requests allow requesting byte ranges of a file
      \item Request from different network interfaces or to different endpoints
        to create alternative paths
      \end{itemize}
  \end{itemize}
